A studyforrest extension, an annotation of spoken language in the German dubbed movie ``Forrest Gump'' and its audio-description

### 2020-05-20 paper/main.tex
revise rest of annotation procedure
revise of methods until 1st lvl analysis

% natural luanguage processing (NLP)
% Tiger Treebank Corpus:
% \href{https://www.ims.uni-stuttgart.de}{ims.uni-stuttgart.de}.
To automatically analyze linguistic features of each words in their corresponding
sentence, we employed the Python package spaCy \citep{spacy2} and its accompanying German language model (\texttt{de\_core\_news\_md} that was trained on the TIGER Treebank corpus \citep{brants2004tiger}.
Non-speech vocalizations were dropped from the sentences before analysis.
We performed analyses regarding part-of-speech (i.e. grammatical tagging or word-category disambiguation), syntactic dependencies, lemmatization, word embedding (i.e. multi-dimensional meaning representation of a word), and if the word is one of the most common words of the German language (i.e. if the word is part of a stop list).

\subsection*{Data Legend}
\todo[inline]{Minimale Info hier, dass es zwei Versionen gibt und sich der folgende Abschnitt auf TSV file bezieht}
The annotation is available in two different versions, both providing the same information: a) as a (text-based) Praat TextGrid file and b) as a text-based, tab-separated value (TSV) formatted table.
The following descriptions refer to the ten columns of the TSV file, namely
\texttt{onset}, \texttt{duration}, \texttt{person}, \texttt{text} (of sentence, word or phoneme), \texttt{pos}, \texttt{tag}, \texttt{dep}, \texttt{lemma}, \texttt{stop}, \texttt{vector}.

@@

\subsubsection*{Speaker identity (\texttt{person})}
Name of the person that speaks the sentence, word or phoneme. See Table \ref{tab:speakers} for the ten most often ocurring speakers.

\subsubsection*{Text (\texttt{text})}
\todo[inline]{Großphoneme erwähnen;Nicht erwähnt: Wenn die Korrektur der Phoneme unmöglich war, habe ich einfach die Event-Borders zwischen den Phonemen herausgenommen. Wortbeginn und Wortende stimmen dann mit einem neuen, "künstlich-großen Phoneme" überein.
Das "Groß-Phonem" enthält dann ohne Grenzen zwischen den tatsächlichen Phonemen des Wortes alle Phoneme des Wortes ("es ist kompliziert zu erklären, aber im Grunde ganz einfach). kommt jedoch so selten vor, ist im Grunde "noise" und ist in der Anno sofort als solches zu erkennen.}
The text of a spoken sentence or word, or the pronounciation of a phoneme.
Phonemes of German words follow the Prosodylab PhoneSet, English words follow the ARPAbet PhoneSet.

\subsubsection*{Simple part-of-speech tag (\texttt{pos})}
A simple part-of-speech tagging (grammatical tagging; word-category disambiguation) of a word.
The simple part-of-speech tagging tags follow the Universal Dependencies v2 POS tag set (\href{https://universaldependencies.org}{universaldependencies.org}).
See Table \ref{tab:pos} for a description of the labels and the respective counts of all 15 labels.
Nouns that spaCy mistook for proper nouns or vice versa were corrected via script.
% s. list at the beginning of script
% code/add_part-of-speech-tagging2textgrid.py
Additionally in cells of this column, sentences are tagged as \texttt{SENTENCE}, and phonemes are tagged as \texttt{PHONEME} to facilitate sorting, and filtering in potential further processing steps.

@@

\subsubsection*{Detailed part-of-speech tag (\texttt{tag})}
% \href{https://www.ims.uni-stuttgart.de/forschung/ressourcen/lexika/germantagsets}{Stuttgart-Tübingen-Tagset}
A detailed part-of-speech tagging of a word following the TIGER Treebank annotation scheme \citep{brants2004tiger} which is based on the Stuttgart-Tübingen-Tagset \citep{schiller1999stts}.
See Table \ref{tab:tag} for a descripion  of the labels and the respective counts of the 15 most often occuring labels. Nouns that spaCy mistook for proper nouns or vice versa were corrected via script.
% s. list at the beginning of script
% code/add_part-of-speech-tagging2textgrid.py

@@

\subsubsection*{Syntactic dependency (\texttt{dep})}
Information about a word's syntactic dependencies with other words within the same sentence.
Information follows the TIGER Treebank annotation scheme \citep{brants2004tiger} and is given in the format:
``arc label;word's head;word's child1, word's child2, ...'', where the ``arc label'' (see Table \ref{tab:dep}) describes the type of syntactic relation that connects a "child" (the single word) to its "head".

@@

\section*{Dataset validation}
% Information about any validation carried out and/or any limitations of the
% datasets, including any allowances made for controlling bias or unwanted
% sources of variability.
% ~170 words => das wird beim folgenden Schreiben definitiv nichts werden
% Intro
In order to assess data quality, we investigated if contrasting speech-related events to events without speech lead to increased activation in areas known to be involved in semantic processing \citep{hickok2007cortical}.
Moreover, we tested if similar linguistic concepts providing high semantic information (proper nouns and nouns) contrasted to a concept providing low semantic information (coordinate conjunctions) lead to increased activation in congruent brain areas.
% Preprocessing
\todo[inline]{hanke2014 spricht von 20 subjects (age 21–38 years, mean age 26.6 years, 12 male), es fehlt hier aber sub-10. MIH: age doesn't matter here. Just say that you left out sub-10, and why}
% Details of the alignment procedure are given in the section on “Technical Validation”.
%Data aligned by applying an affine transformation only are available in bold_dico_dico7Tad2grpbold7Tad.nii.gz.
% Data aligned by non-linear warping are available in bold_dico_dico7Tad2grpbold7Tad_nl.nii.gz.
We used blood oxygenation level-dependent (BOLD) functional magnetic resonance imaging (fMRI) data of the 2h audio-description (\unit[7]{Tesla}, \unit[2]{s} repetition time, 3599 volumes, 36 axial slices, thickness \unit[1.4]{mm}, \unit[1.4 $\times$ 1.4]{mm} in-plane resolution, \unit[224]{mm} field-of-view) of 19 healthy subjects \citep{hanke2014audiomovie}.
Data were already corrected for motion at the scanner computer.
Further, individual BOLD time-series were already aligned by non-linear warping to a study-specific T2*-weighted echo planar imaging (EPI) group template (cf. \citep{hanke2014audiomovie} for details on how the group template was created).
\todo[inline]{So korrekt und “vollständig? MNI152 co-registration? Im Übrigen finde ich 3-4 Paper, die als Zitation für den MNI152-space in Frage kommen, wobei keines richtig passen will; hanke2014 zitiert keines. MIH: That is sufficient. No need to cite anything for MNI152}
All further analysis steps were carried out using Feat v6.00 (FMRI Expert Analysis Tool)\citep{woolrich2001autocorr} as part of FSL v5.0.9 (FMRIB's Software Library)\citep{smith2004fsl}.
Data were temporally high-pass filtered (cut-off \unit[150]{s}), spatially smoothed (Gaussian kernel; \unit[4.0]{mm} FWHM), and the brain was extracted from surrounding tissue.
A grand-mean intensity normalization of the entire 4D dataset was performed by a single multiplicative factor.

@@

The respective counts are given for the whole stimulus and the eight segments that were used during fMRI scanning.
The 20 most often occuring labels from the detailed part-of speech tagging (\texttt{tag}) were used as such.

@@

The label \texttt{sentence} contains the end of complete grammatical sentences. The label \texttt{phones} contains events of the 80 most often occuring phonemes (phoneme \texttt{n} with N=6053 to phoneme \texttt{IY1} with N=32).
The label \texttt{no-sp} represents moments when no speech was audible. \texttt{fg\_ad\_lrdiff} (left-right volume difference) and \texttt{fg\_ad\_rms} (root mean square volume) represent one event for every movie frame (\unit[40]{ms}). Events were convolved with FSL's ``Double-Gamma HRF'' to create the regressors. The correlation of these regressors over the time course of the whole stimulus can be seen in Figure \ref{fig:reg-corr}.}

@@

phones & 80 most often occuring phonemes & \rPhonesAll & \rPhonesI & \rPhonesII & \rPhonesIII & \rPhonesIV & \rPhonesV & \rPhonesVI & \rPhonesVII & \rPhonesVIII \tabularnewline

>>>>

% natural language processing (NLP)
% spaCy
To automatically analyze linguistic features of each words in their
corresponding sentence, we employed the Python package spaCy \citep{spacy2} and
its accompanying German language model (\texttt{de\_core\_news\_md} that was
trained on the TIGER Treebank corpus \citep{brants2004tiger}.
% Tiger Treebank Corpus:
% \href{https://www.ims.uni-stuttgart.de}{ims.uni-stuttgart.de}.
% dropping of non-speech vocalizations because they often delude the algo
Non-speech vocalizations were dropped from the sentences before analysis to
improve results.
% performed analyses
We performed analyses regarding part-of-speech (i.e. grammatical tagging or
word-category disambiguation), syntactic dependencies, lemmatization, word
embedding (i.e. multi-dimensional meaning representation of a word), and if the
word is one of the most common words of the German language (i.e. if the word is
part of a stop list).

\subsection*{Data Legend} \todo[inline]{Minimale Info hier, dass es zwei
Versionen gibt und sich der folgende Abschnitt auf TSV file bezieht}
% TextGrid vs. TSV
The annotation is available in two different versions, both providing the same
information: a) as a (text-based) Praat TextGrid file and b) as a text-based,
tab-separated value (TSV) formatted table.
% preview of following subsubsections
The following descriptions refer to the ten columns of the TSV file, namely
\texttt{onset}, \texttt{duration}, \texttt{person}, \texttt{text} (of sentence,
word or phoneme), \texttt{pos}, \texttt{tag}, \texttt{dep}, \texttt{lemma},
\texttt{stop}, \texttt{vector}.

@@

\subsubsection*{Speaker identity (\texttt{person})} Name of the person that
speaks the sentence, word or phoneme. See Table \ref{tab:speakers} for the ten
most often occurring speakers.

\subsubsection*{Text (\texttt{text})}
% ``Großphoneme'': n<200
The text of a spoken sentence or word, or the pronunciation of a phoneme.
Phonemes of German words follow the Prosodylab PhoneSet, English words follow
the ARPAbet PhoneSet.

\subsubsection*{Simple part-of-speech tag (\texttt{pos})} A simple
part-of-speech tagging (grammatical tagging; word-category disambiguation) of a
word.
% annotation scheme
The simple part-of-speech tagging tags follow the Universal Dependencies v2 POS
tag set (\href{https://universaldependencies.org}{universaldependencies.org}).
% reference to table
See Table \ref{tab:pos} for a description of the labels and the respective
counts of all 15 labels. Nouns that spaCy mistook for proper nouns or vice
versa were corrected via script.
% s. list at the beginning of script code/add_part-of-speech-tagging2textgrid.py
% flags for sentences and phonemes in POS column
Additionally in cells of this column, sentences are tagged as \texttt{SENTENCE},
and phonemes are tagged as \texttt{PHONEME} to facilitate sorting, and filtering
in potential further processing steps.

@@

\subsubsection*{Detailed part-of-speech tag (\texttt{tag})}
% \href{https://www.ims.uni-stuttgart.de/forschung/ressourcen/lexika/germantagsets}{Stuttgart-Tübingen-Tagset}
A detailed part-of-speech tagging of a word following the TIGER Treebank
annotation scheme \citep{brants2004tiger} which is based on the
Stuttgart-Tübingen-Tagset \citep{schiller1999stts}.
% reference to table
See Table \ref{tab:tag} for a descripion  of the labels and the respective
counts of the 15 most often occuring labels. Nouns that spaCy mistook for proper
nouns or vice versa were corrected via script.
% s. list at the beginning of script code/add_part-of-speech-tagging2textgrid.py

@@

\subsubsection*{Syntactic dependency (\texttt{dep})} Information about a word's
syntactic dependencies with other words within the same sentence.
% annotation scheme
Information follows the TIGER Treebank annotation scheme \citep{brants2004tiger}
and is given in the format: ``arc label;word's head;word's child1, word's
child2, ...'', where the ``arc label'' (see Table \ref{tab:dep}) describes the
type of syntactic relation that connects a "child" (the single word) to its
"head".

@@

\section*{Dataset validation}
% Information about any validation carried out and/or any limitations of the
% datasets, including any allowances made for controlling bias or unwanted
% sources of variability. ~170 words => das wird beim folgenden Schreiben
% definitiv nichts werden intro
In order to assess data quality, we investigated if contrasting speech-related
events to events without speech lead to increased activation in areas known to
be involved in semantic processing \citep{hickok2007cortical}.
% nouns and proper nouns resp. > coordinate junctions
Moreover, we tested if similar linguistic concepts providing high semantic
information (proper nouns and nouns) contrasted to a concept providing low
semantic information (coordinate conjunctions) lead to increased activation in
congruent brain areas.
% data source
% data aligned by applying an affine transformation only are available in
% bold_dico_dico7Tad2grpbold7Tad.nii.gz.
% data aligned by non-linear warping are available in
% bold_dico_dico7Tad2grpbold7Tad_nl.nii.gz.
We used a dataset providing blood oxygenation level-dependent (BOLD) functional
magnetic resonance imaging (fMRI) data of 20 subjects (age 21–38 years, mean age
26.6 years, 12 male) listening to the 2h audio-description (\unit[7]{Tesla},
\unit[2]{s} repetition time, 3599 volumes, 36 axial slices, thickness
\unit[1.4]{mm}, \unit[1.4 $\times$ 1.4]{mm} in-plane resolution, \unit[224]{mm}
field-of-view) \citep{hanke2014audiomovie}.
% motion correction
Data were already corrected for motion at the scanner computer.
% alignment
Further, individual BOLD time-series were already aligned by non-linear warping
to a study-specific T2*-weighted echo planar imaging (EPI) group template (cf.
\citep{hanke2014audiomovie} for details on how the group template was created).
% steps done by C.H.
All further analysis steps were carried out using Feat v6.00 (FMRI Expert
Analysis Tool)\citep{woolrich2001autocorr} as part of FSL v5.0.9 (FMRIB's
Software Library)\citep{smith2004fsl}.
% dropping of subject 10
Data of one participant was dropped to due to invalid distortion correction
during scanning.
% temporal and spatial filter, brain extraction
Data were temporally high-pass filtered (cut-off \unit[150]{s}), spatially
smoothed (Gaussian kernel; \unit[4.0]{mm} FWHM), and the brain was extracted
from surrounding tissue.
% normalization
A grand-mean intensity normalization of the entire 4D dataset was performed by a
single multiplicative factor.

@@

The respective counts are given for the whole stimulus and the eight segments
that were used during fMRI scanning.
The 20 most often occurring labels from the detailed part-of speech tagging     (\texttt{tag}) were used as such.

@@

The label \texttt{sentence} contains the end of complete grammatical sentences.
The label \texttt{phones} contains events of the 80 most often occurring
phonemes (phoneme \texttt{n} with N=6053 to phoneme \texttt{IY1} with N=32).
The label \texttt{no-sp} represents moments when no speech was audible. \texttt{fg\_ad\_lrdiff} (left-right volume difference) and \texttt{fg\_ad\_rms} (root mean square volume) represent one event for every movie frame (\unit[40]{ms}).
Events were convolved with FSL's ``Double-Gamma HRF'' to create the regressors.
The correlation of these regressors over the time course of the whole stimulus
can be seen in Figure \ref{fig:reg-corr}.}

@@

phones & 80 most often occurring phonemes & \rPhonesAll & \rPhonesI & \rPhonesII & \rPhonesIII & \rPhonesIV & \rPhonesV & \rPhonesVI & \rPhonesVII & \rPhonesVIII \tabularnewline
